---
title: "Descriptive Visualizations"
author: "Ling Yi & Lunbei Hu"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

# Descriptive visualizations of confirmed COVID cases and deaths in the U.S.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
require(scales)
library(lubridate)
library(httr)
library(patchwork)
library(COVID19)

knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

*Importing relevent dataset which includes the data from the COVID19 R package and a locally made dataset with dates of mask mandate per state.*

```{r message=FALSE, warning=FALSE}
covid_df = covid19(c("US"), level = 2, start = "2020-01-01")

covid_df = 
  covid_df %>% 
  mutate(
    state = key_alpha_2
  ) %>% 
  arrange(state) %>% 
  group_by(state) %>%
  mutate(
    new_case = confirmed - lag(confirmed, default = 0),
    new_case_updated = replace(new_case, which(new_case < 0), NA),
    new_deaths = deaths - lag(deaths, default = 0),
    new_death_updated = replace(new_deaths, which(new_deaths < 0), NA),
    date = as.Date(date),
    new_date = date,
    state_binary = ifelse(state %in% c("ID", "MT", "ND", "WY", "SD", "UT", "NE", "IA", "KS", "MO", "OK", "AR", "TX", "IL", "MS", "AL", "FL", "TN", "KY", "IN", "OH", "WV", "NC", "SC", "AK"), "red", "blue"),
    temperate = ifelse(state %in% c("FL", "TX", "NC", "SC", "GA", "LA", "HI", "OK", "AK", "AL", "MS", "TN", "KY"), "sub-tropical", "temperate")
  ) %>% 
  separate(new_date, into = c("year", "month", "day"))

mask_df = read_excel("./data/mask_state.xlsx") %>% 
  janitor::clean_names() %>% 
  mutate(date = as.Date(as.numeric(as.character(date)), origin = "1899-12-30")) %>%
  dplyr::select(-x4)

cactus_df = left_join(covid_df, mask_df, by.x = c("state", "date"), by.y = c("state", "date"))

```

# State Aggregated New COVID Cases and Deaths


**Comparing new COVID cases by state over the past year^1^**

```{r message=FALSE, warning=FALSE}
time_plot = 
covid_df %>% 
  ggplot(aes(x = date, y = new_case_updated, color = state)) +
  geom_point(alpha = 0.5) +
  geom_line() +
  theme(legend.position = "none") +
  labs(title = "Daily New Covid Cases Over Time",
       y = "Confirmed new cases",
       x = "Date") +
      scale_x_date(breaks = seq.Date(from = as.Date("2020-01-31"), 
                                     to = as.Date("2020-11-01"), by = 15)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5))

time_plot <- time_plot + scale_y_continuous(labels = comma)
time_plot
```

**Comparing new COVID deaths by state over the past year^1^**

```{r message=FALSE, warning=FALSE}
death_plot = 
covid_df %>% 
  ggplot(aes(x = date, y = new_death_updated, color = state)) +
  geom_point(alpha = 0.5) +
  geom_line() +
  theme(legend.position = "none") +
  labs(title = "Daily New Covid Deaths Over Time",
       y = "Confirmed new deaths",
       x = "Date") +
      scale_x_date(breaks = seq.Date(from = as.Date("2020-01-31"), 
                                     to = as.Date("2020-11-01"), by = 15)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5))

death_plot <- death_plot + scale_y_continuous(labels = comma)
death_plot
```

**Daily new COVID case of red/blue state (based on 2020 presidential election) stratified by temperate/sub-tropical state**

```{r message=FALSE, warning=FALSE}
state_plot = 
covid_df %>% 
  group_by(state_binary, date, temperate) %>% 
  summarize(mean_cases = mean(new_case_updated, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean_cases, color = state_binary)) +
  geom_point(alpha = 0.5) +
  geom_line() +
  scale_y_continuous(labels = comma) +
  facet_wrap(~temperate, nrow = 1, scales = "free_x", shrink = FALSE, strip.position = "bottom") +
  labs(title = "Average Daily New COVID cases by Red/Blue states",
       y = "Confirmed new cases",
       x = "Date") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y",expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5))
state_plot
```


**Daily new COVID deaths of red/blue state (based on 2020 presidential election) stratified by temperate/sub-tropical state**

```{r message=FALSE, warning=FALSE}
state_death = 
covid_df %>% 
  group_by(state_binary, date, temperate) %>% 
  summarize(mean_deaths = mean(new_death_updated, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean_deaths, color = state_binary)) +
  geom_point(alpha = 0.5) +
  geom_line() +
  scale_y_continuous(labels = comma) +
  facet_wrap(~temperate, nrow = 1, scales = "free_x", shrink = FALSE, strip.position = "bottom") +
  labs(title = "Average Daily New COVID deaths by Red/Blue states",
       y = "Confirmed new deaths",
       x = "Date") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y",expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5))
state_death
```


```{r message=FALSE, warning=FALSE}
#creating new variables to look at two week difference in new COVID cases and COVID deaths two weeks to one month after mask mandate respectively. 

cactus_df = 
  cactus_df %>% 
  mutate(in_two_weeks_mask = if_else(mask == "Yes", lead(new_case_updated, 14), NA_real_)) %>%
  mutate(diff = in_two_weeks_mask - new_case_updated)


cactus_df = 
  cactus_df %>% 
  mutate(in_two_weeks_death = if_else(mask == "Yes", lead(new_death_updated, 30), NA_real_)) %>% 
  mutate(diff_death = in_two_weeks_death - new_death_updated)
```

**Observing two week difference in confirmed new COVID cases from date of mask mandate in states that enforced a mandate**

```{r message=FALSE, warning=FALSE}
cactus_df %>% 
  ggplot(aes(x = state, y = diff, color = state_binary)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  facet_wrap(~temperate, nrow = 1, scales = "free_x", shrink = TRUE, strip.position =     "bottom") +
 scale_x_discrete(guide = guide_axis(check.overlap = TRUE)) +##not sure if I will use this
    labs(title = "Difference in new cases and mask mandate",
       y = "Two week difference in confirmed new cases",
       x = "State")
```


**Observing two week difference in confirmed new COVID deaths from date of mask mandate in states that enforced a mandate**

```{r message=FALSE, warning=FALSE}
cactus_df %>% 
  ggplot(aes(x = state, y = diff_death, color = state_binary)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  facet_wrap(~temperate, nrow = 1, scales = "free_x", shrink = TRUE, strip.position =     "bottom") +
 scale_x_discrete(guide = guide_axis(check.overlap = TRUE)) +##not sure if I will use this
    labs(title = "Difference in new deaths and mask mandate",
       y = "Four week difference in confirmed new deaths",
       x = "State")
```

# City/County Aggregated New COVID Cases and Deaths




**A comparison of new COVID cases and deaths between main US cities (New York City, Los Angeles, Chicago/Cook County and Houston) and less populated counties**

```{r message=FALSE, warning=FALSE}
confirmed_df = 
  read_csv("https://github.com/CSSEGISandData/COVID-19/raw/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv") %>% 
  select_all(tolower) %>% 
  mutate(
    state = province_state,
    county = admin2
  ) %>% 
  select(-c(uid:combined_key)) %>% 
  select(state, county, everything()) %>% 
  filter(state %in% c("New York", "California", "Illinois", "Texas")) %>% 
  pivot_longer(
    -c(state, county),
    names_to = "date",
    values_to = "case_total"
  ) %>%
  group_by(state, county) %>% 
  mutate(
    date = mdy(date),
    new_case = case_total - lag(case_total, default = 0),
    new_case_updated = replace(new_case, which(new_case < 0), NA)
  )


deaths_df = read_csv("https://github.com/CSSEGISandData/COVID-19/raw/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv") %>% 
  select_all(tolower) %>% 
  mutate(
    state = province_state,
    county = admin2
  ) %>% 
  select(-c(uid:combined_key)) %>% 
  select(state, county, everything()) %>% 
  filter(state %in% c("New York", "California", "Illinois", "Texas")) %>% 
  pivot_longer(
    -c(state, county, population),
    names_to = "date",
    values_to = "death_total"
  ) %>% 
  group_by(state, county) %>% 
  mutate(
    date = mdy(date),
    new_death = death_total - lag(death_total, default = 0),
    new_death_updated = replace(new_death, which(new_death < 0), NA)
  )

confirmed_deaths_df = 
  inner_join(confirmed_df, deaths_df, by = c("state", "county", "date")) %>% 
  group_by(state, county)

lessppl_list =  
  confirmed_deaths_df %>%
  select(state, county) %>% 
  group_by(state) %>% 
  filter(!county %in% c("New York", "Los Angeles", "Cook", "Houston")) %>%
  distinct(county, .keep_all = T) %>%
  sample_n(10)

lessppl_df = 
  inner_join(lessppl_list, confirmed_deaths_df, by = "county") %>% 
  mutate(
    state = state.x
  ) %>% 
  select(-c(state.x, state.y)) %>% 
  select(state, everything())
```



**The most populated city - New York City vs. other less populated counties in New York state regarding new COVID cases and deaths^2^.**

```{r message=FALSE, warning=FALSE}
nyc_df =
  confirmed_deaths_df %>% 
  filter(county == "New York")

ny_lessppl_df = 
  lessppl_df %>% 
  filter(state == "New York") 

ny_df = 
  nyc_df %>% 
  rbind(ny_lessppl_df) %>% 
  mutate(ppl = ifelse(county == "New York", "NYC", "other NY counties"))
```

```{r message=FALSE, warning=FALSE}
ny_confirmed_plot =
  ny_df %>% 
  group_by(ppl, date) %>% 
  summarize(mean_cases = mean(new_case_updated, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean_cases, color = ppl)) +
  geom_point(alpha = 0.5) +
  geom_line() +
  scale_y_continuous(labels = comma) +
  labs(title = "COVID Cases and Deaths of New York City, NY",
       y = "Confirmed new cases",
       x = "Date") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y",expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5))

ny_deaths_plot =
  ny_df %>% 
  group_by(ppl, date) %>% 
  summarize(mean_deaths = mean(new_death_updated, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean_deaths, color = ppl)) +
  geom_point(alpha = 0.5) +
  geom_line() +
  scale_y_continuous(labels = comma) +
  labs(y = "Confirmed new deaths",
       x = "Date") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y",expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5))

ny_confirmed_plot + ny_deaths_plot
```

* NYC has a population of `r format(mean(pull(nyc_df, population)), scientific=FALSE)`, and the mean population of the 10 randomly selected NY counties is `r format(mean(pull(ny_lessppl_df, population)), scientific=FALSE)`.



**The second populated city - Los Angeles vs. other less populated counties in California regarding new COVID cases and deaths^2^.**

```{r message=FALSE, warning=FALSE}
la_df =
  confirmed_deaths_df %>% 
  filter(county == "Los Angeles")

ca_lessppl_df = 
  lessppl_df %>% 
  filter(state == "California") 

ca_df = 
  la_df %>% 
  rbind(ca_lessppl_df) %>% 
  mutate(ppl = ifelse(county == "Los Angeles", "LA", "other CA counties"))
```

```{r message=FALSE, warning=FALSE}
ca_confirmed_plot =
  ca_df %>% 
  group_by(ppl, date) %>% 
  summarize(mean_cases = mean(new_case_updated, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean_cases, color = ppl)) +
  geom_point(alpha = 0.5) +
  geom_line() +
  scale_y_continuous(labels = comma) +
  labs(title = "COVID Cases and Deaths of Los Angeles, CA",
       y = "Confirmed new cases",
       x = "Date") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y",expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5))

ca_deaths_plot =
  ca_df %>% 
  group_by(ppl, date) %>% 
  summarize(mean_deaths = mean(new_death_updated, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean_deaths, color = ppl)) +
  geom_point(alpha = 0.5) +
  geom_line() +
  scale_y_continuous(labels = comma) +
  labs(y = "Confirmed new deaths",
       x = "Date") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y",expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5))

ca_confirmed_plot + ca_deaths_plot
```

* Los Angeles has a population of `r format(mean(pull(la_df, population)), scientific=FALSE)`, and the mean population of the 10 randomly selected CA counties is `r format(mean(pull(ca_lessppl_df, population)), scientific=FALSE)`.


**The third populated city - Chicago (part of Cook County) vs. other less populated counties in Illinois regarding new COVID cases and deaths^2^.**

```{r message=FALSE, warning=FALSE}
chicago_df =
  confirmed_deaths_df %>% 
  filter(county == "Cook")

il_lessppl_df = 
  lessppl_df %>% 
  filter(state == "Illinois") 

il_df = 
  chicago_df %>% 
  rbind(il_lessppl_df) %>% 
  mutate(ppl = ifelse(county == "Cook", "Cook", "other IL counties"))
```

```{r message=FALSE, warning=FALSE}
il_confirmed_plot =
  il_df %>% 
  group_by(ppl, date) %>% 
  summarize(mean_cases = mean(new_case_updated, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean_cases, color = ppl)) +
  geom_point(alpha = 0.5) +
  geom_line() +
  scale_y_continuous(labels = comma) +
  labs(title = "COVID Cases and Deaths of Cook County/Chicago, IL",
       y = "Confirmed new cases",
       x = "Date") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y",expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5))

il_deaths_plot =
  il_df %>% 
  group_by(ppl, date) %>% 
  summarize(mean_deaths = mean(new_death_updated, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean_deaths, color = ppl)) +
  geom_point(alpha = 0.5) +
  geom_line() +
  scale_y_continuous(labels = comma) +
  labs(y = "Confirmed new deaths",
       x = "Date") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y",expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5))

il_confirmed_plot + il_deaths_plot
```

* Cook County (Chicago included) has a population of `r format(mean(pull(chicago_df, population)), scientific=FALSE)`, and the mean population of the 10 randomly selected IL counties is `r format(mean(pull(il_lessppl_df, population)), scientific=FALSE)`.


**The fourth populated city - Houston vs. other less populated counties in Texas regarding new COVID cases and deaths^2^.**

```{r message=FALSE, warning=FALSE}
houston_df =
  confirmed_deaths_df %>% 
  filter(county == "Houston")

tx_lessppl_df = 
  lessppl_df %>% 
  filter(state == "Texas") 

tx_df = 
  houston_df %>% 
  rbind(tx_lessppl_df) %>% 
  mutate(ppl = ifelse(county == "Houston", "Houston", "other TX counties"))
```

```{r message=FALSE, warning=FALSE}
tx_confirmed_plot =
  tx_df %>% 
  group_by(ppl, date) %>% 
  summarize(mean_cases = mean(new_case_updated, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean_cases, color = ppl)) +
  geom_point(alpha = 0.5) +
  geom_line() +
  scale_color_manual(values = c("Houston" = "black", "other TX counties" = "orange")) +
  scale_y_continuous(labels = comma) +
  labs(title = "COVID Cases and Deaths of Houston, TX",
       y = "Confirmed new cases",
       x = "Date") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y",expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) 

tx_deaths_plot =
  tx_df %>% 
  group_by(ppl, date) %>% 
  summarize(mean_deaths = mean(new_death_updated, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean_deaths, color = ppl)) +
  geom_point(alpha = 0.5) +
  geom_line() +
  scale_color_manual(values = c("Houston" = "black", "other TX counties" = "orange")) +
  scale_y_continuous(labels = comma) +
  labs(y = "Confirmed new deaths",
       x = "Date") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b-%y",expand = c(0,0)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5))

tx_confirmed_plot + tx_deaths_plot
```

* Houston has a population of `r format(mean(pull(houston_df, population)), scientific=FALSE)`, and the mean population of the 10 randomly selected TX counties is `r format(mean(pull(tx_lessppl_df, population)), scientific=FALSE)`.


*Texas is one of the sub-tropical states compared to NY, CA and IL (temperate), and it is marked by different colors.*


# National New COVID Cases and Deaths


**Comparing overall US new COVID cases and deaths with important national events**

```{r message=FALSE, warning=FALSE}
library(sars2pack)

us_df = jhu_data() %>% 
  janitor::clean_names() %>% 
  filter(country_region == "US")


us_covid_df = 
  us_df %>% 
    pivot_wider(names_from = subset,
              values_from = count) %>% 
    mutate(
    new_case = confirmed - lag(confirmed, default = 0),
    new_deaths = deaths - lag(deaths, default = 0),
    date = as.Date(date)
    )

events_df = 
  read_excel("./data/national_events.xlsx") 


us_events_df = left_join(us_covid_df, events_df, by.x = c("date"), by.y = c( "date"))


us_events_df = 
  us_events_df %>% 
  mutate(in_two_weeks_mask = if_else(!(is.na(event)), lead(new_case, 14), NA_real_)) %>%
  mutate(diff = in_two_weeks_mask - new_case)


us_events_df = 
  us_events_df %>% 
  mutate(in_two_weeks_death = if_else(!(is.na(event)), lead(new_deaths, 30), NA_real_)) %>% 
  mutate(diff_death = in_two_weeks_death - new_deaths)

```

**Daily new COVID cases over the past year with important national events^3^**

```{r message=FALSE, warning=FALSE}
us_events_df %>%  
  ggplot(aes(x = date, y = new_case)) +
  geom_point() +
    labs(title = "New COVID cases over time",
       y = "Confirmed new cases",
       x = "Date") +
      geom_vline(data = events_df, mapping = aes(xintercept = date) ,linetype = "dotted",color = "blue", show.legend = TRUE) +
  geom_text(data = events_df, mapping = aes(x = date, y = 75000, label = event), size = 2.5, angle = 90, vjust = -0.4, hjust = 0)
```

**Daily new COVID deaths over the past year with important national events^3^**

```{r message=FALSE, warning=FALSE}
us_events_df %>%  
  ggplot(aes(x = date, y = new_deaths)) +
  geom_point() +
    labs(title = "New COVID deaths over time",
       y = "Confirmed new deaths",
       x = "Date") +
      geom_vline(data = events_df, mapping = aes(xintercept = date) , linetype = "dotted", color = "blue", show.legend = TRUE) +
  geom_text(data = events_df, mapping = aes(x = date, y = 1000, label = event), size = 2.5, angle = 90, vjust = -0.4, hjust = 0)
  
```

### Citations

* 1. Guidotti, E., Ardia, D., (2020), "COVID-19 Data Hub", Journal of Open Source Software 5(51):2376, doi: 10.21105/joss.02376.
  
* 2. Dong E, Du H, Gardner L. An interactive web-based dashboard to track COVID-19 in real time. Lancet Inf Dis. 20(5):533-534. doi: 10.1016/S1473-3099(20)30120-1.

* 3. https://github.com/seandavi/sars2pack.
